Adversarial Training of Neural Networks\\ against Systematic Uncertainty

### 2016-08-02 paper.tex
save
method

\glnote{Distinction between statistical and systematic uncertainty.}
\glnote{Define nuisance parameters.}
\glnote{We want to build an accurate classifier whose output remains invariant with
respect to systematic uncertainties.}
% https://www.slac.stanford.edu/econf/C030908/papers/TUAT004.pdf

@@

Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
Let consider the multivariate random variables $X_\lambda: \Omega \mapsto
\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a
nuisance parameter $\lambda$ whose values define the family of its systematic uncertainties.
That is, $X_\lambda$ and $Y$ induce together
a joint probability distribution written as $p(X,Y|\lambda)$, where the
conditional on $\lambda$ denotes $X_\lambda$. For training, let further assume a
finite set $\{ \lambda_i, x_i, y_i, \}_{i=1}^N$ of realizations
$X_{\lambda_i}(\omega_i), Y(\omega_i)$, for $\omega_i \in \Omega$ and known
values $\lambda_i$ of the nuisance parameter. Our goal is to learn a function $f :
\mathbb{R}^p \mapsto {\cal Y}$ (e.g., a classifier if ${\cal Y}$ is a finite set
of classes) minimizing the expected value of a loss $L(Y, f(X_{\lambda}))$, with
the constraint that $f(X_\lambda)$ should be robust to the value of the nuisance
parameter $\lambda$ -- which remains unknown at test time. More specifically, we
aim at building $f$ such that in the ideal case \begin{equation}\label{eqn1}
f(X_{\lambda_i}(\omega)) = f(X_{\lambda_j}(\omega)) \end{equation} for any
sample $\omega \in \Omega$ and any $\lambda_i, \lambda_j$ pair of values of the
nuisance parameter.

Since we do not have training tuples $(X_{\lambda_i}(\omega),
X_{\lambda_j}(\omega))$ (for the same unknown $\omega$), we propose instead to
solve the closely related problem of finding a predictive function $f$ such that
\begin{equation}\label{eqn2}
    P(\{ \omega | f(X_{\lambda_i}(\omega)) = y \} ) = P( \{ \omega' | f(X_{\lambda_j}(\omega')) = y\}) \text{ for all $y \in {\cal Y}$}.
\end{equation}
In words, we are looking for a predictive function $f$ such that  the
distribution of $f(X_\lambda)$ is invariant with respect to the nuisance
parameter $\lambda$. Note that a function $f$ for which Eqn.~\ref{eqn1} is
true necessarily satisfies Eqn.~\ref{eqn2}. The converse is however in
general not true, since the sets of samples $\{ \omega | f(X_{\lambda_i}(\omega)) = y \}$
and $\{ \omega' | f(X_{\lambda_j}(\omega')) = y \}$ do not need to be the same
for the equality to hold. \glnote{This criterion is still adequate for most purposes.}

@@

>>>>

\glnote{Distinction between statistical and systematic uncertainty.}
\glnote{Define nuisance parameters.}
\glnote{We want to build an accurate classifier whose output remains invariant with
respect to systematic uncertainties.}
% https://www.slac.stanford.edu/econf/C030908/papers/TUAT004.pdf
\glnote{Motivate the criterion: allow to derive guarantees no matter $\lambda$,
while $f$ may otherwise be good or very bad depending on the nuisance if trained
on the mixture. E.g. to form hypothesis to be later confirmed by data.}

@@

Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
Let consider the multivariate random variables $X_\lambda: \Omega \mapsto
\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a
nuisance parameter $\lambda$ whose values define the family of its systematic uncertainties.
That is, $X_\lambda$ and $Y$ induce together
a joint probability distribution $p(X,Y|\lambda)$, where the
conditional on $\lambda$ denotes $X_\lambda$. For training, let further assume a
finite set $\{ x_i, y_i, \lambda_i \}_{i=1}^N$ of realizations
$X_{\lambda_i}(\omega_i), Y(\omega_i)$, for $\omega_i \in \Omega$ and known
values $\lambda_i$ of the nuisance parameter. Our goal is to learn a function $f :
\mathbb{R}^p \mapsto {\cal Y}$ (e.g., a classifier if ${\cal Y}$ is a finite set
of classes) minimizing the expected value of a loss $L(Y, f(X_{\lambda}))$, with
the constraint that $f(X_\lambda)$ should be robust to the value of the nuisance
parameter $\lambda$ -- which remains unknown at test time. More specifically, we
aim at building $f$ such that in the ideal case
\begin{equation}\label{eqn:criterion-true}
f(X_{\lambda_i}(\omega)) = f(X_{\lambda_j}(\omega))
\end{equation} for any
sample $\omega \in \Omega$ and any $\lambda_i, \lambda_j$ pair of values of the
nuisance parameter.

Since we do not have training tuples $(X_{\lambda_i}(\omega),
X_{\lambda_j}(\omega))$ (for the same unknown $\omega$), we propose instead to
solve the closely related problem of finding a predictive function $f$ such that
\begin{equation}\label{eqn:criterion}
    P(\{ \omega | f(X_{\lambda_i}(\omega)) = y \} ) = P( \{ \omega' | f(X_{\lambda_j}(\omega')) = y\}) \text{ for all $y \in {\cal Y}$}.
\end{equation}
In words, we are looking for a predictive function $f$ such that  the
distribution of $f(X_\lambda)$ is invariant with respect to the nuisance
parameter $\lambda$. Note that a function $f$ for which Eqn.~\ref{eqn:criterion-true} is
true necessarily satisfies Eqn.~\ref{eqn:criterion}. The converse is however in
general not true, since the sets of samples $\{ \omega | f(X_{\lambda_i}(\omega)) = y \}$
and $\{ \omega' | f(X_{\lambda_j}(\omega')) = y \}$ do not need to be the same
for the equality to hold.

@@

Adversarial training was first proposed by \cite{goodfellow2014generative} as a
way to build a generative model capable of producing samples from random noise
$z \sim p_Z$. More specifically, the authors pit a generative model $G$ against
an adversary classifier $D$ whose repelling objective is to recognize real from
generated data. Both models $G$ and $D$ are trained simultaneously, in such a
way that $G$ learns to produce samples that are difficult to identify by $D$,
while $D$ incrementally adapts to changes in $G$. At the equilibrium, $G$ models
a distribution whose samples can be identified by $D$ only by chance. In other
words, assuming enough capacity in $D$ and  $G$, the distribution $p_{G(Z)}$
eventually converges towards the real distribution $p_X$.

In this work, we repurpose adversarial training for regularizing the
construction of the predictive model $f$ so as to satisfy
Eqn.~\ref{eqn:criterion}.

\glnote{describe baseline}
\glnote{describe adversarial approach}
\glnote{proof that it solves Eqn. 2}
